---
output:
  md_document:
    variant: markdown_github
---

### Risk Parity Portfolios (RPPs): A Comparative Study in South Africa

**Overview**  
This repository contains resources and analyses for assessing the applicability of Risk Parity Portfolios (RPPs) in South African markets. The study compares RPPs to Minimum Variance Portfolios (MVPs) and Maximum Diversification Portfolios (MDPs) with a focus on risk-adjusted performance, economic regime sensitivity, and robustness across different market conditions.

**Research Goals**  
1. Evaluate the performance of RPPs relative to MVPs and MDPs.  
2. Investigate RPP viability under South African economic conditions from 2007 to 2024.  
3. Analyze RPPs' adaptability across different economic regimes (high growth, moderate growth, and recession).  

**Core Metrics**  
The study utilizes a comprehensive set of performance measures:  
- Risk-adjusted returns: Sharpe, Sortino, Sterling, and Calmar ratios.  
- Downside risk: Value at Risk (VaR), Conditional VaR, and drawdown metrics.  

**Dataset**  
The portfolio construction uses data on South African stocks and government bond yields (2- and 10-year maturities) covering 2007 to 2024. The dataset spans diverse market conditions, including the Global Financial Crisis and the COVID-19 pandemic.  

**Methodology**  
1. Construct portfolios using RPP, MVP, and MDP methodologies.  
2. Benchmark portfolio performance using downside risk and risk-adjusted return metrics.  
3. Evaluate portfolio performance during distinct economic regimes.  

**Key Findings**  
1. **RPP Strengths**: Effective in minimizing short-term downside risk; robust in downturns.  
2. **RPP Weaknesses**: Underperformance in cumulative and risk-adjusted returns relative to MVPs and MDPs, especially in high-growth conditions.  
3. Economic Regime Analysis: RPPs demonstrated varying performance, with significant challenges during recessions.  

**Conclusion**  
This study highlights the trade-offs inherent in leveraging RPPs for portfolio diversification in South Africa. While RPPs offer robust risk management, their returns are less competitive compared to MVPs and MDPs.  


```{r}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

Loading the necessary libraries and data sets. 
```{r, warning=FALSE}
library(PerformanceAnalytics)
library(riskParityPortfolio)
#library(fPortfolio)
library(dplyr)
library(tidyr)
library(xts)
library(lubridate)
library(kableExtra)
library(RiskPortfolios)
library(PortfolioAnalytics)
library(PerformanceAnalytics)
library(ROI)
library(ROI.plugin.quadprog)
library(zoo)
library(DEoptim)
LCL_Stock_Returns <- read_rds("/Users/x/Downloads/Data_2024/LCL_Stock_Returns.rds")
BondYields_10Y_SA <- read_rds("/Users/x/Downloads/Data_2024/BondYields_10Y_SA.rds")
BondYields_2Y_SA <- read_rds("/Users/x/Downloads/Data_2024/BondYields_2Y_SA.rds")
SA_Growth <- read_csv2("/Users/x/Downloads/SA_Growth.csv", show_col_types = FALSE)
```

In a first step, the filter_stock_function will be applied to the LCL_Stock_Returns dataset. The function filters and transforms stock data by removing weekends, reshaping data into a wide format with tickers as columns, dropping columns with only missing values or constant values, and converting the result into an xts time-series object.

```{r}
LCL_filtered <- filter_stock_function(LCL_Stock_Returns)
```

In the following chunk, the weights and returns for the MVP and the MDP are calculated. The weights for the MVP are calculated using the rolling_min_vol_portfolio_weights function and for the MDP the most_diversified_weights function. 

The MVP function calculates portfolio weights for a rolling minimum volatility strategy. The function optimizes portfolio weights over a specified window of historical returns window_size and adjusts these weights at regular intervals defined by the rebalance_interval. Key constraints ensure the portfolio remains fully invested, uses only long positions, and individual asset weights are bounded between 0 and 1. At each rebalance date, the portfolio optimization uses the most recent window_size days of data, employing a variance minimization objective. The resulting weights are normalized, stored as an xts time-series object, and validated to ensure they are non-negative and sum to one for every rebalancing period.

The MDP function calculates portfolio weights for the MDP strategy. This function optimizes weights to maximize the diversification ratio, which is the ratio of the weighted sum of asset volatilities to the portfolio's overall volatility. The portfolio is rebalanced at regular intervals (rebalance_interval). Asset weights are constrained to be non-negative and bounded between 0 and 1. Optimization is performed using the Differential Evolution (DEoptim) algorithm, with covariance and standard deviation of asset returns used as inputs. 

The returns for each portfolio are then calculated using the Return.portfolio package from the Performance Analytics package.

```{r}
weights_mv <- min_vol_portfolio_weights(LCL_filtered, 252, 63)
returns.mv <- Return.portfolio(LCL_filtered, weights = weights_mv, verbose = TRUE)
weights_md <- most_diversified_weights(LCL_filtered, window_size = 252, rebalance_interval = 63) 
returns.md <- Return.portfolio(LCL_filtered, weights = weights_md, verbose = TRUE)
```

In order to include the bonds in one of the portfolios, the returns of the bonds has to be calculated. THis is done using the Bond_return function. Calculates bond returns based on yield and maturity. The function processes the input data by selecting relevant columns, renaming the yield column, and calculating bond prices using the formula 
$$ \text{BondPrice} = \frac{1}{(1 + \text{Yield}/100)^\text{Maturity}} . $$

```{r}
Bond_return_10 <- Bond_return(BondYields_10Y_SA, "BondYield_10", 10)
Bond_return_2 <- Bond_return(BondYields_2Y_SA, "BondYield_2", 2)
```

After merging the bond and stock returns the covariance matrix is calculated.

```{r}
LCL_Bond <- merge.xts(Bond_return_10, Bond_return_2, LCL_filtered, join = "inner")
Sigma <- cov(LCL_Bond)
```

The riskParityPortfolio function from the same named package will be used to determine the weights for the RPP. This function is embedded in the RollingRiskParity function which rebalances the portfolio every three months. 

```{r}
portfolio.parity <- riskParityPortfolio(Sigma)
rWindows<-rollingWindows(LCL_Bond, period="12m",
                         by="3m")
parity.weights <- RollingRiskParity(rWindows$from@Data, rWindows$to@Data, LCL_Bond) %>% na.omit()
parity.returns <- Return.portfolio(LCL_Bond, weights=parity.weights,verbose=TRUE)
```

The same steps are repeated for the portfolio without bonds. For further analysis all portfolios are then merged to one data frame.

```{r}
Sigma_nb <- cov(LCL_filtered)
portfolio.parity.nb <- riskParityPortfolio(Sigma_nb)
rWindows.nb<-rollingWindows(LCL_filtered, period="12m",
                         by="3m")
parity.weights.nb <- RollingRiskParity(rWindows.nb$from@Data, rWindows.nb$to@Data, LCL_filtered) %>% na.omit()
parity.returns.nb <- Return.portfolio(LCL_filtered, weights=parity.weights.nb,verbose=TRUE)
p.returns.nb <- merge.xts(returns.mv$returns, parity.returns.nb$returns, parity.returns$returns, returns.md$returns, join = "inner")
names(p.returns.nb)<-c("MVP", "RPP w/o Bonds", "RPP", "MDP")
```

For a first impression on the performance, the cumulative returns and the annualized returns are calculated.

```{r}
PerformanceAnalytics::chart.CumReturns(p.returns.nb, colorset=rich6equal,
                          lwd=1, cex.legend = 0.5, event.labels = TRUE, main = "", legend.loc = "topleft")
```

```{r}
chart.RollingPerformance(p.returns.nb, width = 252, colorset=rich6equal,
                          lwd=1, cex.legend = 0.5, event.labels = TRUE, main = "", legend.loc = "top")
```

To quantify the results they are then calculated and the numeric values of both are displayed in a table.

```{r}
bind_rows(
  as.data.frame(Return.cumulative(p.returns.nb)),
  as.data.frame(Return.annualized(p.returns.nb))
) %>% kbl(caption = "Cumulative and Annualized Returns", label = "returns") %>% 
    kable_styling()
```

Since the focus of the portfolios is to minimize risk, the drawdowns are very interesting to visually compare how the portfolios behave when the financial markets are falling.

```{r}
chart.Drawdown(p.returns.nb, colorset=rich6equal,
                          lwd=1, cex.legend = 0.5, event.labels = TRUE, main = "", legend.loc = "bottom")
```

To get more insights, in addition to the chart a table with different drawdown measures such as the Sterling, Calmar, Burke, Pain, and Martin Ratios, along with the Pain Index and the Ulcer Index, will be calculated.

```{r}
table.DrawdownsRatio(p.returns.nb, Rf = 0.08975) %>%
  kbl(caption = "Drawdowns ratio table", label = "drawdowns-ratio") %>%
  kable_styling()
```

To compare the downside risk, different Sharpe Ratios will be calculated.

```{r}
SharpeRatio(p.returns.nb, Rf = 0.095) %>% 
    kbl(caption = "Sharpe Ratios", label = "sharpe-ratios")  %>% 
    kable_styling()
```

The calculate_ratios function filters bond yield data to include only observations after 2007 and determines the risk-free rate as the average 10-year bond yield. It computes the Sortino Ratio for risk-adjusted returns, Conditional Drawdown at Risk (CDD) at a 95% confidence level, the Upside Potential Ratio for assessing return potential relative to downside risk, Conditional Value at Risk (CVaR) for average losses beyond the Value at Risk, and Value at Risk (VaR) for maximum expected losses. The results are returned as a data frame of calculated ratios.

```{r}
calculate_ratios(p.returns.nb, BondYields_10Y_SA) %>% 
    kbl(caption = "Downside and Opportunity-based Performance Metrics", label = "downside-metrics") %>%
    kable_styling()
```

To evaluate the performance of RPPs across various economic environments, their performance will be analyzed during periods of recession, moderate growth, and high growth. The economic_regime function classifies years based on real GDP growth rates, while the regime_return function integrates the output of the economic_regime function and computes three performance metrics: return, volatility, and Sharpe ratio.

```{r}
return_parity <- return_data(parity.returns)
return_mv <- return_data(returns.mv)
return_parity_nb <- return_data(parity.returns.nb)
return_md <- return_data(returns.md)

regime <- economic_regime(SA_Growth, 3, 0)

results_mv <- regime_returns(return_mv, regime)
results_RPP <- regime_returns(return_parity, regime)
results_RPP_nb <- regime_returns(return_parity_nb, regime)
results_md <- regime_returns(return_md, regime)

combine_regime_results(results_mv, results_RPP, results_RPP_nb, results_md) %>% 
    kbl(caption = "Performance at Different Economic States", label = "regime") %>% 
    kable_styling()
```



